---
title: "Household Power Consumption Analysis"
author: "Rubén Guerrero Ramírez"
date: "20 de julio de 2018"
output: 
  html_document:
    toc: True
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
options(scipen = 999)
```

# Introduction

The goal of this document is to report all the steps done through the analysis of a household power consumption. 
The main purpose of the analysis is to understand the behaviour of a family in their home through the use of the energy measured through different submeters. Also, the cost of the energy consumption will be analyzed in order to make a prediction of future energy behaviour.  

# Data Set Description

The data collected is from a french house, probably located in Clamart. It contains 2.075.259 measurements gathered between December 2006 and November 2010. The measurements were recorded with a one-minute sampling rate over 47 months.
The data set can be donwloaded [here](http://archive.ics.uci.edu/ml/datasets/Individual+household+electric+power+consumption).

## Attribute Information

The dataset contains 9 columns described below:

 1. **date**: Date in format `dd/mm/yyyy`.
 2. **time**: Time in format `hh:mm:ss`.
 3. **golbal_active_power**: Household global minute-averaged active power (in kilowatt).
 4. **global_reactive_power**: Household global minute-averaged reactive power (in kilowatt) 
 5. **voltage**: Minute-averaged voltage (in volt) 
 6. **global_intensity**: Household global minute-averaged current intensity (in ampere) 
 7. **sub_metering_1**: Energy sub-metering No. 1 (in watt-hour of active energy). It corresponds to the kitchen, containing mainly a dishwasher, an oven and a microwave (hot plates are not electric but gas powered). 
 8. **sub_metering_2**: Energy sub-metering No. 2 (in watt-hour of active energy). It corresponds to the laundry room,  containing a washing-machine, a tumble-drier, a refrigerator and a light. 
 9. **sub_metering_3**: Energy sub-metering No. 3 (in watt-hour of active energy). It corresponds to an electric water-heater and an air-conditioner.

# Import Data

In firt place, we are going to import the raw data through `read.csv` and assign it into `initialPowerData`. If you open the raw file you will find that the separator between attributes is the semicolon (";").

```{r Initial import}

initialPowerData <- read.csv2("household_power_consumption.txt")

# Alternative
# initialPowerData <- read.csv("household_power_consumption.txt", sep = ";")

```

The next step will be to check what is inside the data we have imported through `str()` and `summary()` functions.

```{r Initial summary}

str(initialPowerData)
summary(initialPowerData)

```

The first thing that shows up is that all variables are codified as factors, we will need to change them to the proper format (`Date` and `Time` will be converted to date type variables and the rest of the variables will be converted to numeric). Also, all variables with energy measurements have 25979 cells with question mark ("?") in each variable, which can be considered as missing values (except for `Sub_metering_3`). Variable `Sub_metering_3` instead of having question mark on missing values has empty values (""). In order to fix this issue we are going to import the data again setting NA's as question marks.
 

```{r proper import}

powerData <- read.csv2("household_power_consumption.txt", na.strings = c("?",""))

```

Let's check again the structure and the summary of the imported data:

```{r proper summary}

summary(powerData)

```

Now we can see on the summary that all cells with question mark or empty space are now readed as missing values NA's.

# Acknowledgements

The data used belongs to:

  Dua, D. and Karra Taniskidou, E. (2017). UCI Machine Learning Repository [http://archive.ics.uci.edu/ml]. Irvine, CA: University of California, School of Information and Computer Science. 